---
title: "TidyTuesday - Plants in Danger"
author: "wjtorres"
date: '2020-08-18'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(ggthemes) # for theme I like
library(ggtext) # to add color to certain variable levels
library(patchwork) # for side by side plots
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-08-18")

```

# Pull data

```{r}

plants <- tt$plants
threats <- tt$threats
actions <- tt$actions

```

# View plants data

```{r}
head(plants)
tail(plants)

# view plants structure
str(plants)

# glimpse
glimpse(plants)

# is each binomial name unique?
length(unique(plants[["binomial_name"]])) 

# which countries are in plants?
n_distinct(plants$country) # there are 72 countries
unique(plants$country) # and here are the 72 country names
plants %>%
  group_by(country) %>%
  summarise(percentage = n()/500) %>%
  arrange(desc(percentage))

# which continents are in plants?
n_distinct(plants$continent) # there are 6 continents
unique(plants$continent) # and here are the 6 continent names

# what groups are in plants?
n_distinct(plants$group) # there are 6 groups
unique(plants$group) # and here are the 6 plant names

# what red list categories are in plants?
n_distinct(plants$red_list_category) # there are 2 groups in this data frame for plants that are considered to be in danger
unique(plants$red_list_category) # the group labels are "Extinct in the wild" or "Extinct"

# How is year_last_seen measured? it looks like it is a character string
n_distinct(plants$year_last_seen)
unique(plants$year_last_seen) # looks like 8 date ranges in increments of 19 years



```

# Data visualization - where are the plants in most danger located?

```{r, fig.width = 6, fig.asp = 0.618}

plot1 <- plants %>%
  group_by(country) %>%
  summarize(n = n()) %>%
  mutate(highlight_flag = ifelse(country %in% c("Madagascar", "United States", "Ecuador"), T, F)) %>%
  ggplot(aes(reorder(country,n), n, fill = highlight_flag)) +
  geom_bar(stat = "identity", show.legend = F) +
  coord_flip() +
  theme_fivethirtyeight() +
  labs(
    title = "As of 2020,there are 500 plant species considered extinct",
    subtitle = "Most are located in Madagascar, United States, and Ecuador",
    caption = "source =  Data from International Union for Conservation of Nature (IUCN) Red list of Threatened Species (Version 2020-1)",
    x = "Country",
    y = "Red List Category"
    ) +
  scale_fill_manual(values = c('#595959', 'darkorange1'))

plot1

```
# Data Visualization - When where the plants last seen for countries with the most extinct plants?

```{r, fig.width = 7, fig.asp = 0.618}

# select only countries with most extinct plants 

plants_small <-
  plants %>%
  filter(country %in% c("Madagascar", "United States", "Ecuador"))

# create ordered variables for graphing

plants_small$year_last_seen_factor_ordered <- factor(plants_small$year_last_seen, ordered = TRUE, 
                                levels = c("Before 1900", "1900-1919", "1920-1939", "1940-1959", "1960-1979", "1980-1999", "2000-2020"))

plants_small$country_ordered <- factor(plants_small$country,levels = c("Ecuador", "United States", "Madagascar"))

# plot

plot2 <- plants_small %>%
  filter(!is.na(year_last_seen_factor_ordered)) %>%
  ggplot(aes(country_ordered, fill = year_last_seen_factor_ordered)) +
  geom_bar() +
  theme_fivethirtyeight() +
  labs(
    title = "When were plants last seen?",
    fill = "Year last seen:"
    ) +
  theme(legend.position = "bottom") +
  coord_flip()

plot2

```

# Data visualization - what are the threats for the plants that are in most danger?

```{r,fig.width = 10, fig.asp = 0.618}

# focus  on countries with most extinct plants

threats_small <- threats %>%
   filter(country %in% c("Madagascar", "United States", "Ecuador"))

# create ordered variables

threats_small$country_ordered <- factor(threats_small$country, levels = c("Ecuador", "United States", "Madagascar"))

# plot

plot3 <- threats_small %>%
  filter(threatened == 1) %>%
  ggplot(aes(country_ordered, fill = threat_type)) +
  geom_bar() +
  theme_fivethirtyeight() +
  labs(
    title = "What are the threats?",
    fill = "Threat type:"
    ) +
  theme(legend.position = "bottom") +
  coord_flip()

plot3
```

# Data visualization - side by side by side
```{r, fig.width = 22, fig.asp = 0.618}

plot1 + (plot2/plot3)

```